\section{Potter's Square Root Covariance filter}

Normal Kalman filters (as described above) work directly with the
covariance of the data, and the dynamic range in magnitudes involved
can cause numerical issues. Some of these issues can be addressed by
instead working directly with a square-root form of the data. Since
the covariance matrix $P$ is positive-definite, we can calculate another
matrix $S$ such that $S S^t = P$ (call this the square root matrix).
Potter's mechanization of the Kalman filter uses the $S$ matrix directly
as part of the filter state, exhibiting improved numerical properties.

This algorithm operates on one scalar observation at a time, and
requires using the normalized ${\bar H}$ matrix and ${\bar z}$ vectors
to eliminate the $R$ matrix as discussed earlier. ${\bar H}_j$ is used
to refer to the $j$-th row of the normalized coefficient matrix, and
${\bar z}_j$ refers to the $j$-th entry in the normalized observation vector.

\subsection{Data Processing}
Given a Potter filter state
representing an estimate of the state of the system,
and an observation vector containing additional information
about the system,
a refined estimate of the state of the system
can be generated using the following equations,
where observations are presented one scalar value at a time,
with
${\hat x}$ and ${\hat S}$
from each iteration
carrying over as
${\tilde x}$ and ${\tilde S}$
for the next iteration:
\begin{subequations}
\label{potterdata}
\begin{align}
\nu & = {\bar H}_j {\tilde S}
\\
\sigma & = 1 / ( \nu \nu^t + 1 )
\\
K & = {\tilde S} \nu^t
  & \mbox{\it (unweighted Kalman Gain)}
\\
{\hat x} & = {\tilde x} +
    \sigma \left( {\bar z}_j - {\bar H}_j {\tilde x} \right) K
  & \mbox{\it (improved State estimate)}
\label{potterdatax}
\\
{\hat S} & = {\tilde S} - \frac{\sigma}{1 + \sqrt{\sigma}} K \nu
  & \mbox{\it (improved Error estimate)}
\label{potterdatas}
\end{align}
\end{subequations}
See \verb|potter_data_c.c|, \verb|potter_data_f.f| and \verb|potter_data.m|
for example implementations.
The symbols $\nu$ and $\sigma$
are used here as convenient names
for local scratch storage,
and have no significance
outside this specific procedure.
